Causal Bandits: Learning Good Interventions via Causal Inference

### 2016-05-04 causal-bandit/nips_2016/simple-regret.tex
started re-writing section 4
updates definition of m for simple regret

\begin{figure}[h]
\centering
\begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1cm,
  thick,main node/.style={observed}, hidden/.style={empty}]

@@

\end{tikzpicture}
\caption{Causal model for the parallel bandits problem.\label{fig:causalStructure}}
\end{figure}

@@

The algorithm operates as follows. For the first $T/2$ rounds it chooses $do()$ to collect observational data.
The fact that the only link from each $X_1,\ldots,X_N$ to $Y$ is a direct, causal one means that $\P{Y|do(X_i=j)}=\P{Y|X_i=j}$. Thus we can create good estimators of the returns of those actions $do(X_i = j)$ for which $\P{X_i = j}$ is large. 
These are the interventions that are easily estimated ``passively'' (\ie, without playing them explicitly). 
Those actions where $\P{X_i = j}$ is small may not be observed (often) in the first $T/2$ rounds, so
the estimates of the returns of these actions from the first $T/2$ samples could be quite poor. For the remaining $T/2$ rounds
the algorithm makes interventions on the actions $a = do(X_i = j)$ for which $\P{X_i = j}$ is small. The precise allocation is
carefully chosen to minimise the worst-case simple regret. 
The difficulty of the problem depends on $\vec{q}$ and, in particular, how many of the variables have an infrequently occurring value (\ie, small $q_i$ or $1-q_i$). 
Let $I_1 = \emptyset$ and for $\tau \in [0,1)$ let $I_\tau = \set{ i : \min\set{q_i, 1-q_i} \leq \tau}$.
Define
\eqn{
\label{eq:m-simple}
m(\vec{q}) = \max \set{ d : |I_{\frac{1}{d}}| \leq d}\,.
% m(\vec{q}) = \max \set{i : \sum_{k=1}^N \ind{\min\set{q_k, \,1 - q_k} \leq 1/i} \leq i}\,.
}
This captures a trade off between identifying the low probability variables and not having too many of them.
When $\vec{q} = (\frac{1}{2}, \ldots, \frac{1}{2})$ we have $m(\vec{q}) = 1$ and when $\vec{q} = (0, \ldots, 0)$ we have $m(\vec{q}) = N$, the number of variables.

\begin{algorithm}[h]
\caption{Parallel Bandit Algorithm}\label{alg:simple}
\begin{algorithmic}[1]
\STATE {\bf Input:} Total rounds $T$ and $N$.
\FOR{$t \in 1,\ldots,T / 2$}
\STATE Perform empty intervention $do()$
\STATE Observe $\vec{X}_t$ and $Y_t$
\ENDFOR
\FOR{$a = do(X_i = x) \in \actions$}
\STATE Count times $X_i = x$ seen: $T_a = \sum_{t=1}^{T/2} \ind{X_{t,i} = x}$
\STATE Estimate reward: $\hat{\mu}_a = \frac{1}{T_a} \sum_{t=1}^{T/2} \ind{X_{t,i} = x} Y_t$ \\[0.2cm]
\STATE Estimate probabilities: $\hat{p}_a = \frac{2 T_a}{T}$,\,\, $\hat q_i = \hat p_{do(X_i = 1)}$
\ENDFOR
\STATE Compute $\hat{m} = m(\vec{\hat q})$ and $A = \set{a \in \actions \colon \hat{p}_a \leq \frac{1}{\hat m}}$.
\STATE Let $T_A := \frac{T}{2 |A|}$ be times to sample each $a\in A$.
\FOR{$a = do(X_i = x) \in A$}
\FOR{$t \in 1,\ldots,T_A$}
\STATE Intervene with $a$ and observe $Y_t$
\ENDFOR
\STATE Re-estimate $\hat{\mu}_a = \frac{1}{T_A} \sum_{t=1}^{T_A} Y_t$
\ENDFOR
\RETURN estimated optimal $\hat{a}^*_T \in \argmax_{a\in\actions} \hat{\mu}_a$
\end{algorithmic}
\end{algorithm}

@@

>>>>

\begin{wrapfigure}[11]{r}{0.5\textwidth}
  \begin{center}
    \begin{tikzpicture}[->,>=stealth',shorten >=1pt,auto,node distance=1cm,
  thick,main node/.style={observed}, hidden/.style={empty}]

@@

\end{tikzpicture}
  \end{center}
  \caption{Causal model for the parallel bandits problem.\label{fig:causalStructure}}
\end{wrapfigure}

@@

The algorithm operates as follows. For the first $T/2$ rounds it chooses $do()$ to collect observational data. As the only link from each $X_1,\ldots,X_N$ to $Y$ is a direct, causal one, $\P{Y|do(X_i=j)}=\P{Y|X_i=j}$. Thus we can create good estimators of the returns of those actions $do(X_i = j)$ for which $\P{X_i = j}$ is large. 
These are the interventions that are easily estimated ``passively'' (\ie, without playing them explicitly). 
Those actions where $\P{X_i = j}$ is small may not be observed (often) in the first $T/2$ rounds, so
the estimates of the returns of these actions from the first $T/2$ samples could be quite poor. For the remaining $T/2$ rounds
the algorithm makes interventions on the actions $a = do(X_i = j)$ for which $\P{X_i = j}$ is small. The precise allocation is
carefully chosen to minimise the worst-case simple regret. 
The difficulty of the problem depends on $\vec{q}$ and, in particular, how many of the variables have an infrequently occurring value (\ie, small $q_i$ or $1-q_i$).

%Let $I_1 = \emptyset$ and for $\tau \in [0,1)$ let $I_\tau = \set{ i : \min\set{q_i, 1-q_i} \leq \tau}$.
For $\tau \in [1...N]$ let $I_\tau = \set{ i : \min\set{q_i, 1-q_i} \leq \frac{1}{\tau - 1}}$.
Define
\eqn{
\label{eq:m-simple}
m(\vec{q}) = \min \set{ d : |I_{d}| \leq d}\,.
% m(\vec{q}) = \max \set{i : \sum_{k=1}^N \ind{\min\set{q_k, \,1 - q_k} \leq 1/i} \leq i}\,.
}

This captures a trade off between identifying the low probability variables and not having too many of them.
When $\vec{q} = (\frac{1}{2}, \ldots, \frac{1}{2})$ we have $m(\vec{q}) = 1$ and when $\vec{q} = (0, \ldots, 0)$ we have $m(\vec{q}) = N$, the number of variables.

@@

\begin{wrapfigure}[22]{r}{0.6\textwidth}
\vspace{-10pt}
\begin{minipage}{.6\textwidth}
\begin{algorithm}[H]
\caption{Parallel Bandit Algorithm}\label{alg:simple}
\begin{algorithmic}[1]
\STATE {\bf Input:} Total rounds $T$ and $N$.
\FOR{$t \in 1,\ldots,T / 2$}
\STATE Perform empty intervention $do()$
\STATE Observe $\vec{X}_t$ and $Y_t$
\ENDFOR
\FOR{$a = do(X_i = x) \in \actions$}
\STATE Count times $X_i = x$ seen: $T_a = \sum_{t=1}^{T/2} \ind{X_{t,i} = x}$
\STATE Estimate reward: $\hat{\mu}_a = \frac{1}{T_a} \sum_{t=1}^{T/2} \ind{X_{t,i} = x} Y_t$ \\[0.2cm]
\STATE Estimate probabilities: $\hat{p}_a = \frac{2 T_a}{T}$,\,\, $\hat q_i = \hat p_{do(X_i = 1)}$
\ENDFOR
\STATE Compute $\hat{m} = m(\vec{\hat q})$ and $A = \set{a \in \actions \colon \hat{p}_a \leq \frac{1}{\hat m}}$.
\STATE Let $T_A := \frac{T}{2 |A|}$ be times to sample each $a\in A$.
\FOR{$a = do(X_i = x) \in A$}
\FOR{$t \in 1,\ldots,T_A$}
\STATE Intervene with $a$ and observe $Y_t$
\ENDFOR
\STATE Re-estimate $\hat{\mu}_a = \frac{1}{T_A} \sum_{t=1}^{T_A} Y_t$
\ENDFOR
\RETURN estimated optimal $\hat{a}^*_T \in \argmax_{a\in\actions} \hat{\mu}_a$
\end{algorithmic}
\end{algorithm}
\end{minipage}
\end{wrapfigure}
